---
title: "PSTAT 126 FINAL PROJECT"
author: "Helen Cui, Becky Su, Sarah Hermann"
date: "3/14/2019"
output:
  github_document:
    math_method:
      engine: webtex
      url: http://chart.apis.google.com/chart?cht=tx&chl=
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r extract data}
library(readxl)
library(car)
data.ccpp<-read_excel("/Users/beckysu/Desktop/ccpp.xlsx")
data(data.ccpp)
attach(data.ccpp)

a<-data.ccpp$AT
b<-data.ccpp$V
c<-data.ccpp$AP
d<-data.ccpp$RH
Y<-data.ccpp$PE
```

```{r plot}
pairs(~Y+a+b+c+d)

full.lm<-lm(Y~a+b+c+d)
avPlots(full.lm, id=FALSE)
```

```{r Residual}
Residualtest<-full.lm$residuals
Fitted<-full.lm$fitted.values
plot(Residualtest~Fitted, xlab = 'Fitted Values', ylab = 'Residuals', main = 'Residuals vs Fits Plot')
abline(h = 0, lty = 2)
```

```{r Q-Q}
e<-Residualtest
qqnorm(e)
qqline(e)
```

```{r Outliers fu}
ccpp.lm <- lm(Y~a+b+c+d)
s.resid <- rstudent(ccpp.lm)
as=abs(s.resid)
which(as==max(as))

#test for outliers
outlierTest(ccpp.lm)
```

```{r influential point fu}
influenceIndexPlot(full.lm)
```

```{r}
library(leaps)
#Full model
mod.0 <- lm(Y~1)
mod.full= ~a + d + b + c

mod.1 <- update(mod.0, mod.full)
mod.backward <- step(mod.1, scope = c(lower = ~ 1, direction = 'backward'))
```

```{r powerTransform}
Trans.ccpp<-powerTransform(cbind(a, b, c, d)~1)
summary(Trans.ccpp)
```

From using power transform, the likelihood ratio tests indicate that using log transformations for all variables is not appropriate, neither should we use no transformations because both p-values ($2.22e^{-16}$) are less than the default alpha level of 0.05. By rounding the estimated lambda powers, 1.2705 to 1, -0.1366 to 0, -2.9843 to -3, and 1.6195 to 1, we decided to log transform the predictor b and not transform the predictor c because it would over complicate our model. 

```{r BoxCox}
ccpp.Trans <- with(data.ccpp, data.frame(Y, a, log(b), c, d))
ccpp.lm<-lm(Y~., data = ccpp.Trans)
boxCox(ccpp.lm)
```

By using box cox transformation, the lambda value is between -1 and -2 so we chose to use lambda=-2 since the confidence level is closer to -2 than -1. Thus our final transformed model with the interaction terms is $Y^{-2}$~a+log(b)+c+d. 

```{r} 
pairs(~I(Y^(-2))+a+log(b)+c+d)

full.fit<-lm(Y^(-2) ~ a + log(b) + c+ d)
avPlots(full.fit, id=FALSE)
```
```{r}
library(leaps)
#Full model
mod.full= ~a +log(b) + c + d + a*b

# Base model
mod.0 <- lm(Y^(-2)~1)
mod.1 <- update(mod.0, mod.full)
mod.backward <- step(mod.1, scope = c(lower = ~ 1, direction = 'backward'))
```

By using backward selection, we found out that all predictors are important predictors. Thus, the final model is $Y^{-2}$ ~ a + log(b) + c + d + a*b.  

```{r new residuals}
Residual<-full.fit$residuals
Fitted<-full.fit$fitted.values
plot(Residual~Fitted, xlab = 'Fitted Values', ylab = 'Residuals', main = 'Residuals vs Fits Plot')
abline(h = 0, lty = 2)
```

```{r new Q-Q}
e<-Residual
qqnorm(e)
qqline(e)
```

```{r Outliers}
ccpp.lm <- lm(Y^(-2)~a+log(b)+c+d+a*b)
s.resid <- rstudent(ccpp.lm)
as=abs(s.resid)
which(as==max(as))

#test for outliers
outlierTest(ccpp.lm)
```

```{r influential point}
full.fit<-lm(Y^(-2) ~ a + log(b) + c + d+a*b)
influenceIndexPlot(full.fit)
```

```{r remove influential points}
ccpp.lm<-lm((Y^(-2) ~ a + log(b) + c + d+a*b), subset=-c(3118,7665,3384,3896,7399,8718,8363,8188,1439,4219))
```

```{r}
Residual<-ccpp.lm$residuals
e<-Residual
qqnorm(e)
qqline(e)
```

```{r global F-test}
#PROBLEM 3B:
ccpp.full.lm<-lm(Y^(-2)~a+log(b)+c+d+a*b)
summary(ccpp.full.lm)
```

The null hypothesis for the global F-test for this model using log(Life) as the response is:
$H_0$: $\beta_1$=$\beta_2$=$\beta_3$=$\beta_4$=$\beta_{(12)}$=0 and the alternative hypothesis is
$H_1$: at least one these predictors is not 0. Since the p-value is $2.2e^{-16}$ which is smaller than the default alpha level of 0.05, we reject the null hypothesis and conclude that at least one predictor is useful.

```{r partial F-test}
#PROBLEM 3D:
ccpp.red.lm<-lm(Y^(-2) ~ a + log(b) + d +a*b)
summary(ccpp.red.lm)
```

The null hypothesis for the partial F-test to test if ambient pressure is an useful predictor is:
$H_0$: $\beta_3$=0 and the alternative hypothesis is
$H_1$: $\beta_3$ is not 0. Since the p-value is $2.2e^{-16}$ which is smaller than the default alpha level of 0.05, we reject the null hypothesis and conclude that ambient pressure is an useful predictor. 

Adjusted $R^2$: 93.21% of the variability of energy output is explained by the model.